---
title: "Wrangling With the Music: An Intro to Data Wrangling"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse) # Automatically loads ggplot2 and dplyr
knitr::opts_chunk$set(echo = FALSE)
spotify_songs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')
```


## Data Wrangling Tutorial

Welcome to our data wrangling tutorial where we will show you a few basic functions to help you wrangle data in R! For this tutorial we will be using a dataset about spotify music, stay tuned!`

### Check out the dataset!

*The dataset was made available by Charlie Thompson, Josiah Parry, Donal Phipps, and Tom Wolff who authored the spotifyr package as well as Kaylin Pavlik who had a recent blogpost using the audio features to explore and classify songs. She used the spotifyr package to collect about 5000 songs from 6 main categories (EDM, Latin, Pop, R&B, Rap, & Rock).*

More information about the dataset can be found here: [TidyTuesday: Spotify Songs](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-21/readme.md).

Feel free to click the link above and view the data dictionary which breaks down the attributes in spotify_songs (e.g. track_name, playlist_genre, energy).

Run this code to view the data on spotify_songs:

```{r view-data, exercise=TRUE, exercise.eval=TRUE}
spotify_songs
```

## Question: Most popular tracks?


How can we use data wrangling to find the most popular tracks on the list? Well first we'll need to learn a few functions in R that will help us out.

### Introducing select(), distinct(), top_n(), and arrange()
The first function we will learn is: select().

#### select()
Select() takes a subset of coulmns that you choose and will display them. So instaead of having to look at all of the columns in this
dataset, we can narrow it down to the ones we want to see. Let's try it out!

First, to see which track is the most popular, we will need to look at these columns: "track_name" and "track_popularity". Let's select these coulmns and call the new variable: "spotify_albums"
```{r select_columns, exercise=TRUE}
spotify_albums = spotify_songs %>% 
  select(track_name, track_popularity) #selects the two columns that we want

spotify_albums
```


Wow, that really makes things  much easier to look at. However, it looks like some of the track names repeat? No worries, theres a way to
fix that. We can get rid of duplicate rows with the distinct() function!

#### distinct()

The distinct() function can be used to keep only unique/distinct rows from a data frame. If there are duplicate rows, only the first row
is preserved.
```{r distinct, exercise=TRUE}
spotify_albums = spotify_songs %>% 
  select(track_name, track_popularity) %>% 
  distinct() #takes the data set and gets rid of any duplicate data

spotify_albums
```

#### top_n()

Now how do we look at the 10 most popular tracks? we use the top_n() function!
```{r topn, exercise=TRUE}
top10 = spotify_songs %>% 
  select(track_name, track_popularity) %>% 
  distinct() %>% 
  top_n(10)

top10
```

#### arrange()

Uh oh, somethings a bit off...I think we need to arrange the data into descending order. We can do this easily with the arrange() function. The arrange() function simply sorts rows in either ascending or descending order:
```{r arrange, exercise=TRUE}
top10 = spotify_songs %>% 
  select(track_name, track_popularity) %>% 
  distinct() %>% 
  top_n(10) %>% 
  arrange(desc(track_popularity)) #arranges the data in descending order

top10
```

## Question: Most popular albums?

Now let's try to answer this next question with the skills we learned so far and with some new tricks!

First, lets select the columns we want to view: "track_album_name" , "track_popularity"
```{r select_again, exercise=TRUE}
spotify_songs %>% 
  select(track_album_name, track_popularity)
```

### Introducing summarize() and groupby()

summarize() and group_by() allow us to make calculations and comparisons.

summarize() (with or without group_by()) collapses the data into rows based on summary statistics.

The important step in summarize() is specifying how we want to reduce the data to a single set of summary statistics.

Examples: mean(), sd(), median(), IQR(), max(), n(), and many more.

Now let's try out the summarize() function to find the mean popularity of all the songs in the dataframe:
```{r summarize_alone, exercise=TRUE}
#summarize on its own as an example
spotify_songs %>% 
  select(track_album_name, track_popularity) %>%
  summarize(mean_pop = mean(track_popularity, na.rm = TRUE))
```

Note: We added na.rm = TRUE to remove NA (missing values) from each calculation.

Otherwise, the output for mean() would also be NA.

#### Using summarize() with group_by().

We just used summarize() to calculate the mean duration of all songs.

But what if we want to calculate these summary statistics for each genre?

We can use group_by() in the pipeline to generate aggregate summary statistics!

The summarize() function is almost always used in conjunction with group_by().

Run the R code below to view the mean popularity of songs per album with `groupby()`:
```{r summarize_groupby, exercise=TRUE}
#using summarize and group_by
spotify_songs %>% 
  select(track_album_name, track_popularity) %>%
  group_by(track_album_name)
  summarize(mean_pop = mean(track_popularity, na.rm = TRUE))

spotify_songs
```

We added the step group_by(track_album_name), which does exactly what it sounds like: it groups by track_album_name before calculating summary statistics!

Note: We never use the group_by() operation by itself. We always combine it with a function that performs a calculation, such as summarize().


Now that we have learned how to use summarize and group_by, let's try to answer our question of: what are the most popular albums? 

Using the new skills that we learned with summarize and goup_by, try to complete the code to calculate the mean:
*hint: what should we group by and what are we trying to summarize?
```{r finding_mean, exercise=TRUE}
spotify_songs %>% 
  select(track_album_name, track_popularity) %>%
  group_by( ) %>% 
  summarize(mean_pop = (track_popularity, na.rm = TRUE)) %>% 
  arrange(desc(track_name))

spotify_songs
 
```

Now let's see the top 5 popular albums using top_n():

```{r top5, exercise=TRUE}
spotify_songs %>% 
  select(track_album_name, track_popularity) %>%
  group_by(track_album_name) %>% 
  summarize(mean_pop = mean(track_popularity, na.rm = TRUE)) %>% 
  arrange(desc(track_album_name)) %>% 
  top_n(5)

```

Let's see how this would look as a graphic visual:
```{r graphic, exercise=TRUE}
top_albums %>% 
  ggplot(aes(x = track_album_name, y = mean_pop)) +
  geom_col() +
  coord_flip() +
  labs(title = "5 Most Popular albums on Spotify", x = "Album Name", y ="Average Rating")
```

While this graph doesn't seem like much, it shows a visual of the most popular albums on spotify based on the mean ratings of each song per album. It should also be noted that some "albums" are also singles and therefore this data shows both actual albums and singles.
## Example: Avg Danceability


Using the spotify_songs data frame as well as the functions `summarize()` and `mean()`, let's calculate the average danceability for all of the songs:

```{r mean_danceability, exercise=TRUE}
spotify_songs %>%
  summarize(mean_danceability = mean(danceability, na.rm = TRUE))
```


## Example: Avg Danceability per Playlist Genre

Now let's see how we can use the functions we just learned to find out more about our data!

Run the R code below to view the mean danceability of each playlist_genre using `groupby()`:

```{r avg_genre_danceability, exercise=TRUE}
avg_genre_danceability = spotify_songs %>%
  select(playlist_genre, danceability) %>% 
  group_by(playlist_genre) %>% 
  summarize(mean_danceability = mean(danceability, na.rm = TRUE)) %>%
  arrange(mean_danceability)

avg_genre_danceability
```

First, we selected the playlist_genre and danceability columns, since we will need that data to answer our question! Then, we grouped our data by the playlist_genre and used summarize() to show the average danceability of each playlist_genre. Lastly, we used arrange() to sort the data by the least to the most danceable song! 


## Exercise: Avg Energy per Playlist Subgenre

Now let's apply some of the data wrangling techniques we have learned!


Using the spotify_songs data frame, calculate the `mean` `energy` level of each `playlist_subgenre`. Assign the pipeline to the name `avg_subgenre_energy`. If you get stuck, feel free to scroll back through the tutorial for help!
```{r avg_subgenre_energy, exercise=TRUE, exercise.lines = 7}

```

---
correct answer:
avg_subgenre_energy = spotify_songs %>%
  select(playlist_subgenre, energy) %>% 
  group_by(playlist_subgenre) %>% 
  summarize(mean_energy = mean(energy, na.rm = TRUE)) %>%
  arrange(mean_energy)

avg_subgenre_energy
---

## Quiz

According to the results in the previous exercise, answer the question below:


```{r quiz}
quiz(
  question("Which subgenre has the least energy?",
    answer("pop"),
    answer("permanent wave"),
    answer("neo soul", correct = TRUE),
    answer("tropical")
  )
)
```


