---
title: "Data Wrangling Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse) # Automatically loads ggplot2 and dplyr
knitr::opts_chunk$set(echo = FALSE)
spotify_songs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')
```


## Intro to Data Wrangling

### Wrangling Spotify Music Data

Tutorial intro: Insert description of data wrangling and dataset

### Check out the dataset!

Run this code to view the data on spotify songs!

```{r view-data, exercise=TRUE, exercise.eval=TRUE}
spotify_songs
```


## Richney's Part

## Question: Most popular tracks?

How can we use data wrangling to find the most popular track on the list? Well first we'll need to learn a few functions in R that will help us out.

The first function we will learn is: select()
select() takes a subset of coulmns that you choose and will display them. So instaead of having to look at all of the columns in this dataset, we can narrow it down to the ones we want to see. Let's try it out!

First, to see which track is the most popular, we will need to look at these columns: "track_name" and "track_popularity"
```{r}
popular_track = spotify_songs %>% 
  select(track_name, track_popularity)

popular_track
```


Wow, that really makes things  much easier to look at. However, it looks like some of the track names repeat? No worries, theres a way to fix that. We can get rid of duplicate rows with the distinctt() function!
```{r}
popular_tracks = popular_track %>% distinct()
popular_tracks
```

Now how do we look at the 10 most popular tracks? we use the top_n() function!
```{r}
top_10 = popular_tracks %>% top_n(10)
top_10
```

Uh oh, somethings a bit off...i think we need to arrange() the data into descending order:
```{r}
top_10 %>% arrange(desc(track_popularity))
```

## Question: Most popular albums?

Now let's try to answer this next question with the skills we learned so far and with some new tricks!
First, lets select the coulmns we want to view.
```{r}
spotify_songs %>% 
  select(track_album_name, track_name, track_popularity)
```

Introducing summarize() and groupby()

summarize() and group_by() allow us to make calculations and comparisons.

summarize() (with or without group_by()) collapses the data into rows based on summary statistics.

The important step in summarize() is specifying how we want to reduce the data to a single set of summary statistics.

Examples: mean(), sd(), median(), IQR(), max(), n(), and many more

Now let's try out the summarize() function to find the mean popularity of all the songs in the dataframe:
```{r}
#summarize on its own as an example
spotify_songs %>%
  summarize(mean_pop = mean(track_popularity, na.rm = TRUE))
```

Note: We added na.rm = TRUE to remove NA (missing values) from each calculation.

Otherwise, the output for mean() would also be NA.

### Using summarize() with group_by().

We just used summarize() to calculate the mean duration of all songs.

But what if we want to calculate these summary statistics for each genre?

We can use group_by() in the pipeline to generate aggregate summary statistics!

The summarize() function is almost always used in conjunction with group_by().

Run the R code below to view the mean danceability of each playlist_genre using `groupby()`:
```{r}
#using summarize and group_by
spotify_songs %>% 
  group_by(track_album_name) %>% 
  summarize(track_name = n()) %>% 
  arrange(desc(track_name))
```

We added the step group_by(track_album_name), which does exactly what it sounds like: it groups by track_album_name before calculating summary statistics!

Note: We never use the group_by() operation by itself. We always combine it with a function that performs a calculation, such as summarize().


Now that we have learned how to use summarize and group_by, let's try to answer our question of: what are the most popular albums? How?

We're gonna use the spotify_songs data frame,and mean ratings per album!

Summarize mean ratings per album:
```{r}
popular_albums = spotify_songs %>%
  group_by(track_album_name) %>% 
  summarize(track_name = n(),
            mean_pop = mean(track_popularity, na.rm = TRUE)) %>% 
  arrange(desc(mean_pop))
```

lumping factors together for a graph:
```{r}

```



## Kaya's Part: summarize() and groupby()


Using the spotify_songs data frame as well as the functions `summarize()` and `mean()`, let's calculate the average danceability for all of the songs:

```{r mean_danceability, exercise=TRUE}
spotify_songs %>%
  summarize(mean_danceability = mean(danceability, na.rm = TRUE))
```


### Example: Average Danceability per Playlist Genre

Now let's see how we can use the functions we just learned to find out more about our data!

Run the R code below to view the mean danceability of each playlist_genre using `groupby()`:

```{r avg_genre_danceability, exercise=TRUE}
avg_genre_danceability = spotify_songs %>%
  select(playlist_genre, danceability) %>% 
  group_by(playlist_genre) %>% 
  summarize(mean_danceability = mean(danceability, na.rm = TRUE)) %>%
  arrange(mean_danceability)

avg_genre_danceability
```

First, we selected the playlist_genre and danceability columns, since we will need that data to answer our question! Then, we grouped our data by the playlist_genre and used summarize() to show the average danceability of each playlist_genre. Lastly, we used arrange() to sort the data by the least to the most danceable song!


### Exercise: Average Energy per Playlist Subgenre

Now let's apply some of the data wrangling techniques we have learned!

Using the spotify_songs data frame, calculate the `mean` energy level `energy` of each `playlist_subgenre`. Assign the pipeline to the name `avg_subgenre_energy`.


```{r avg_subgenre_energy, exercise=TRUE, exercise.lines = 7}

```

---
correct answer:
avg_subgenre_energy = spotify_songs %>%
  select(playlist_subgenre, energy) %>% 
  group_by(playlist_subgenre) %>% 
  summarize(mean_energy = mean(energy, na.rm = TRUE)) %>%
  arrange(mean_energy)

avg_subgenre_energy
---

### Quiz

According to the results in the previous exercise, answer the question below:


```{r quiz}
quiz(
  question("Which subgenre has the least energy?",
    answer("pop"),
    answer("permanent wave"),
    answer("neo soul", correct = TRUE),
    answer("tropical")
  )
)
```


